Edge sign prediction in general graphs and Correlation Clustering

### 2017-09-19 troll/experiments.tex
add a new set of troll experiments

We now evaluate our edge sign classification methods on representative real-world datasets of
varying density and label regularity. This shows that our methods compete well against
existing approaches in terms of both predictive and computational performance. We are especially
interested in small training set regimes, and have restricted our comparison to the batch learning
scenario since all competing methods we are aware of have been developed in that setting only.

@@

>>>>

We now evaluate our edge sign classification methods on representative real-world datasets of
varying density and label regularity.
After presenting the data and our evaluation criterion, we proceed in two steps. First, we simulate
our generative model on real networks to give them signs, then we study to which extent we can recover
the parameters $p$ and $q$ of each node, and how this translates to accurate prediction. Second, we
select random training sets from the actual signs.
This shows that our methods compete well against
existing approaches in terms of both predictive and computational performance. We are especially
interested in small training set regimes, and have restricted our comparison to the batch learning
scenario since all competing methods we are aware of have been developed in that setting only.

@@

\begin{newcontent}
\paragraph{Synthetic signs}

Recall that according to our generative model of \autoref{s:gen}, each node $u$ has parameter $p_u$
governing its sending behavior and each node $v$ has a parameter $q_v$ governing its receiving
behavior such that the sign of the edge \euv{} is positive with probability $\frac{p_u+q_v}{2} =
\etauv$. Given a topology $G=(V,E)$, we start by assigning a $p$ and $q$ values to each node. If we
want \etauv{} to be uniform over $E$, we have to take into account the out-degree of nodes with at
least one outgoing edges, and likewise the in-degree of nodes with at least one incoming edges. For
that, we partition the interval $[0,1]$ into $|V|$ segments of size proportional to $\dout(u_1),
\dout(u_2), \ldots, \dout(u_{|V|})$. We shuffle these segments and draw a number \uar{} from each of
them, which we set as the $p$ value of the corresponding node. Then, because we want to have more
positive edges, we apply an exponential transformation \[ p \rightarrow \frac{1}{1-e^{-\lambda}}
\left( 1-e^{-\lambda p} \right) \,,\] where we choose $\lambda=3$. We do the same for $q$, then for
every edge \euv{}, we set it positive with probability \etauv{}. An example of the distributions we
obtain for the \wik{} graph is showed on \autoref{fig:troll_synth_distrib}.

Once the signs are generated, we select a training set \trainset{} \uar{}, and we want to predict
the signs of the remaining edges in the testing set $E\setminus \trainset{}$. The Bayes optimal
predictor classifies an edge \euv{} as positive if $\etauv \geq \shalf$ and negative otherwise. 
We compare it with the algorithm analyzed in \autoref{ss:bayes_approx}, which we call \usrule{}
(Bayes Learning Classifier based on \emph{tr}ollness and \emph{un}trustworthiness). In practice,
\usrule{} proceeds as follow: after computing $\htr(u)$ and $\hun(u)$ on training set $E_0$ for all
$u \in V$ (or setting those values to \shalf{} in case there is no outgoing or incoming edges for
some node), we use the equation \eqref{eq:predictor}, having estimated $\tau$ on $E_0$.

The results of this comparison on the six datasets are showed in \autoref{tab:troll_synth_global}.
For each network, we generated the signs once. Then for different training size (20\%, 40\% or 80\%
of $E$), we sampled a training set and predict using the knowledge of the true $\eta$ (Bayes
predictor) or its estimated value (\usrule{}).
The MCC of the Bayes predictor is the same on all datasets (around $30$), and so is its accuracy
(around $75$). The corresponding values for \usrule{} are close on all datasets (although some are
easier) and the gap naturally decreases as the training size increases. Another interesting
quantity is how close can \usrule{} estimates $\eta$ using $\wh{\eta} = \left(1-\htr(u)\right) +
\left(1-\hun(v)\right) -\tauhat$ as defined in \autoref{ss:bayes_approx}. We compute the mean
absolute error (MAE) between $\eta$ and $\wh{\eta}$ for all testing edges, and specifically for
the testing edges whose both endpoints have been sampled above a certain threshold in the training
set. As expected, the estimation is more accurate with increasing training set size, and with
increasing number of samples for a given edge.

\input{synth_res}

Finally, we can also look in more details at the role of $p$ and $q$ by building a 2D histogram of
the testing edges \euv{} based on their coordinate $(p_u, q_v)$. For instance on
\autoref{fig:tr_synth_mae}, we see that the MAE between $\eta$ and $\wh{\eta}$ is not uniform over
the $p,q$-space. It seems to decrease along the diagonal from $(0,0)$ to $(1,1)$. Note though that
\autoref{fig:tr_synth_sizemae} suggests this is likely due to the higher number of edges in the top
right corner (reflecting the imbalance toward positive edges), which allows better estimation.

In this case, we cannot use the MCC because off-diagonal, the Bayes predictor classifies edges
either all positive or all negative, resulting in a division by zero in the definition. We thus
fall back on accuracy and display in \autoref{fig:tr_synth_acc} the difference between the accuracy
of the Bayes predictor and the accuracy of \usrule{}. This time, the anti diagonal where
$\frac{p+q}{2}=\lhalf$ seems to play a special role. Indeed, as showed more clearly in
\autoref{fig:tr_synth_cmpacc}, the gap in accuracy between the Bayes predictor and \usrule{}
increases symmetrically as $\frac{p+q}{2}$ moves away from \shalf{}. The gap also tends to disappear
when $\frac{p+q}{2}=\lhalf$, most likely because in that area, prediction can only be made by
chance.

While we show those patterns on \wik{} with a 40\% training size, they are consistent across
datasets and training size.

\begin{figure*}[hbtp]
  % \centering
  \begin{subfigure}[b]{0.504\textwidth} 
    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_mae.pdf} 
    \caption{Mean absolute error (MAE) of the \etauv{} estimation.} 
    \label{fig:tr_synth_mae}
  \end{subfigure}~
  \begin{subfigure}[b]{0.504\textwidth} 
    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_sizemae.pdf} 
    \caption{MAE in each cell as a function of the number of sampled edges.} 
    \label{fig:tr_synth_sizemae}
  \end{subfigure}

  \begin{subfigure}[b]{0.504\textwidth} 
    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_acc.pdf} 
    \caption{100 times the accuracy.} 
    \label{fig:tr_synth_acc}
  \end{subfigure}~
  \begin{subfigure}[b]{0.504\textwidth} 
    \includegraphics[width=\textwidth]{raw/synthgen_wik_rule_40_cmpacc.pdf} 
    \caption{Difference of accuracy between \usrule{} and Bayes as $\frac{p+q}{2}$ moves away from $\frac{1}{2}$.} 
    \label{fig:tr_synth_cmpacc}
  \end{subfigure}
  \caption{Results on \wik{} with \usrule{} on a training set of size 40\%}
\end{figure*}
